Finally Tagless, Partially Evaluated

### 2008-09-16 tagless/tagless-final.tex
rephrase a sentence to ensure that it is simple to check that it is grammatically correct

Our contribution to the literature on partial evaluation is to use mere
Hindley-Milner types in the metalanguage to assure statically and patently that
partially evaluating a well-typed object program not only never gets stuck but
also, if it terminates, produces a well-typed output program that never gets
stuck.  Moreover, these types form an instance of a general |Symantics|
signature that encompasses other interpreters such as evaluation and
compilation.  This early and manifest assurance of type safety contrasts, for
example, with \citearound{'s compiler generator (cogen) for
ML}\citet{Birkedal-PE-ML}, which transforms a program into its tagless
generating extension.  Because that cogen uses a universal type, the fact that
it never generates an ill-typed generating extension from a well-typed input
program is only manifest when each generating extension is type-checked, and the
fact that the generating extension never generates an ill-typed residual program
from well-typed static input is only manifest when each residual program is
type-checked.  Similarly, the fact that \citet{fiore:nbe-ppdp2002} and
\citets{balat:tdpe-popl2004} partial evaluator using delimited control
operators never turns well-typed code into ill-typed code is not assured by the
metalanguage, whether or not as part of a typed family of interpreter modules.
\begin{comment}
% Removed the following comment in view of Section 4.5 of Danvy's
% original TDPE paper (POPL 1996)
It is
type-directed, so the user must represent, as a term, the type of every
term to be partially evaluated.  We shift this work to the type checker
of the metalanguage.
By avoiding term-level type representations, our
approach makes it easier to perform algebraic simplifications (as
in~\S\ref{S:PE-solution}).
\end{comment}

>>>>

Our contribution to the literature on partial evaluation is to use mere
Hindley-Milner types in the metalanguage to assure statically and patently that
partially evaluating a well-typed object program not only never gets stuck but
also, if it terminates, produces a well-typed output program that never gets
stuck.  Moreover, these types form an instance of a general |Symantics|
signature that encompasses other interpreters such as evaluation and
compilation.  This early and manifest assurance of type safety contrasts, for
example, with \citearound{'s compiler generator (cogen) for
ML}\citet{Birkedal-PE-ML}, which transforms a program into its tagless
generating extension.  Because that cogen uses a universal type, the fact that
it never generates an ill-typed generating extension from a well-typed input
program is only manifest when each generating extension is type-checked, and the
fact that the generating extension never generates an ill-typed residual program
from well-typed static input is only manifest when each residual program is
type-checked.  Similarly, the fact that the partial evaluator of
\citet{fiore:nbe-ppdp2002}, nor that of
\citet{balat:tdpe-popl2004}, both of which use delimited control
operators, never turn well-typed code into ill-typed code is not assured by the
metalanguage, whether or not as part of a typed family of interpreter modules.
\begin{comment}
% Removed the following comment in view of Section 4.5 of Danvy's
% original TDPE paper (POPL 1996)
It is
type-directed, so the user must represent, as a term, the type of every
term to be partially evaluated.  We shift this work to the type checker
of the metalanguage.
By avoiding term-level type representations, our
approach makes it easier to perform algebraic simplifications (as
in~\S\ref{S:PE-solution}).
\end{comment}
